\documentclass[14pt,a4paper]{memoir}
\usepackage[margin=0.1in]{geometry}

\begin{document}

\chapter*{First Singing Lesson with James Eisner}

\begin{itemize}
\item Balance evenly on both feet, and between forward and back

\item Back of neck straight not tilted back (crushes larynx)

\item Drop Jaw (which I think was another way of saying don't tilt head back rather than an instruction to do anything with jaw itself)

\item Shield and Ring / Thyroid and Cricoid cartilages

\item Singing is the opposite of swallowing. When you swallow, everything closes up, you want it all as open as possible.
So can practice swallowing and learn what it feels like to relax everything.

\item Drinking water is good to keep everything lubricated

\item Cat and Cow exercise, on all fours arch and hollow back: inhale to midpoint and exhale to both extremes.
makes you conscious of how your breathing apparatus is working and the feeling of things filling up with air

\item Back Breathe: hang off door handle, tuck in tailbone, let go of lower back, feel expansion under ribs at back
expansion goes into pelvis front and back and side. 

\item Reflexive Breathe: sink down and cross arms while exhaling, stand up and allow breathing in to happen automatically.

While making a note, tap all inflating areas (breastbone, ribs, abdomen, sides, lower back) and feel effect in voicebox (if connected, note should rise)
\item When doing this at James, all areas produced effect, originally could not reproduce at home, but back breathing exercise above brings it all back by getting more bits involved in singing.

\item Lower spine is over-concave, stand straight, get taller
Alexander technique imagine your suspended from the ceiling by a string to the top of your head
If you get this right then it improves the connection to the various abdomen areas as tested above.

\item Sing on 'ng' to put all air through nasopharynx

\item Eurythmy (actions as physical mnemonics for various notes)

\item Use the French time names, but h is better than s for rests, should be able to find them plus recordings on internet

\item Metrical Hierarchy (stress beat 1, and also 3 when counting in fours)

\end{itemize}

\end{document}
